### 2009-09-29 trunk/latex/paper1/contents.tex
corrigindo erro nas regras de classificao do modelo LF.

% TODO: CITACAO DEPOIS DO CONCEITO
In the context of our research, we consider classes as the components

\section{Abstract}

Software clustering algorithms automatically recognize a system's
modular structure by analyzing its implementation.
Due to the lack of well document software systems, though, the issue of testing
these algorithms is still underexplored, limiting both their adoption in the
industry and the development of better algorithms.
We propose to rely on software models to produce arbitrarily large test sets. In
this paper we consider three such models and analyze how similar the artifacts
they produce are from artifacts from real software systems.

@@

One of the most obvious ways to reduce the time needed to develop or maintain a
software system is to assign more developers to it. When combined with poor
management, though, adding developers may in fact raise the development time,
due to the increasing communication cost between developers \cite{Brooks1995}.
Poor management may result in developers working on overlapping tasks, writing
duplicated code, and inadvertently breaking the code written by another
developer.

%The development and maintenance of large software systems is a big challenge for
%software engineering \cite{Northrop2006}. In order to reduce the development
%time of such a system, a tempting idea is to assign more developers for the
%project.

In order to take advantage of a team of developers, therefore, a key issue is
the ability to decompose a system into weakly coupled modules, so that each
module is maintained by a distinct group of developers \cite{Parnas1972}.
Because of the weak coupling between modules, the need for communication between
groups of developers is reduced, and so is the development time.

The ability of decomposing a system into modules depends on global knowledge
about how its implementation components (e.g., functions, classes, variables)
interact. Often, specially in legacy systems, this information is not readily
available. Instead, it must be extracted from the system's source code. Even
then the task of finding modules may be overwhelming because of the large number
of components and interactions.

The task can be made easier by the use of software clustering algorithms, also
known as also known as architecture recovery algorithms \cite{Pollet2007}. They
automatically decompose a software system into weakly coupled modules by
analyzing the network of dependencies between its components.

% sw clustering is a task of reverse engineering \cite{Tonella2007}.

An important question for the adoption of a software clustering algorithm is
whether the decomposition it finds for a system is similar to a reference
decomposition for that system, i.e., a decomposition found by a group of
experienced developers. To answer this question one needs to test the algorithm
by applying it to a sample of systems with known reference decompositions, and
then compare the two decompositions for each system by means of a metric such as
MoJo \cite{Tzerpos1999} and EdgeSim \cite{Mitchell2001}.

Unfortunately there are few systems with known reference decompositions.
Furthermore, because such decompositions are costly do obtain
\cite{Koschke2000}, there are few empirical studies about software clustering
algorithms, most of them based on a couple of small and medium systems
\cite{Anquetil1999,Maqbool2007,Bittencourt2009}.

@@

In this research we propose to evaluate software clustering algorithms by
applying them to synthetic, i.e., computer generated, component dependency
networks with built-in decompositions. With this approach it is feasible to test
algorithms with arbitrarily large samples in a controlled manner.

Of course, it is desirable that the synthetic networks resemble networks
extracted from real software systems. Hence, in this paper, we describe a study
about three models that synthesize networks with built-in decompositions. We
show empirically that all three models are capable of synthesizing networks that
resemble software networks.

The remaining sections are organized as follows. 
Section \ref{sec:swnets}...
Section \ref{sec:models}...
Section \ref{sec:characterization}...
Section \ref{sec:evaluation}...
Section \ref{sec:conclusion}...

%alike, resembling, exchangeable, indistiguible

%\section{Related Work}
%
%\section{Software Modularization Recovery: an Overview}
%Aka software architecture recovery, software architecture reconstruction,
%software clustering...
%It's an task of reverse engineering, as defined by Tonella \cite{Tonella2007}...
%DEFINITIONS
%external edge...

%\section{Software Networks and Software Clustering}
%
%directed graph, (un)weighted

\section{Software Networks} \label{sec:swnets}
% TODO: consider merging with introduction

At the implementation level, a software system can be viewed as a network of
interacting components. In a object-oriented system, for example, classes
interact with other classes through mechanisms such as inheritance and
aggregation. By abstracting the particular mechanisms of interaction, it is
possible to build a network of dependencies between components, represented as a
directed graph. This kind of network is the input for many software clustering
algorithms \cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005}.

In this paper we will study networks of dependencies between classes in
object-oriented systems. We will call any network of this kind a \emph{software
network}. This representation of a software system is amenable to analysis
through network theory. %Software networks can be studied by network theory.

%Define: Component. Component dependency network. Directed unweighted graph.
%Clustering based on dependencies.

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, it has been found that many networks
arising from sociology, biology, technology and other domains present remarkable
structural similarities. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{\-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to software networks and it was shown that they
are also scale-free networks \cite{Myers2003,Valverde2003}. Thus, the study of
the structure of software systems in general can be benefited from the vast
literature available about scale-free networks.

%(in fact we can't argue the degree distribution is perfectly fit by a power law.
%Anyway, the distribution is much more assymetric than normal ou Poisson's)
%
%software indegree distribution (power law), outdegree distribution (not so power
%law)

@@

Many models were proposed to explain the formation of scale-free networks. These
models are algorithms that can be proven, either formally or empirically, to
synthesize networks that are scale-free. Most models, though, do not account for
modular structure. In this section we present three models that synthesize
directed networks with built-in modular decompositions: BCR+, CGW and LF.

%The first two models grow networks using a mechanism called
%preferential attachment CITE Barabasi, by which nodes with many edges tend to
%receive more edges in the process.

%\subsection{ER (Erdos-Renyi)}

%\subsection{Configuration model}
% take the most typical software system and steal its degree sequence

@@

The algorithm 1R from data mining was used to help discover such relationship.
It analyzes the parameters and the classification of each network and finds a
rule that relates the value of one single parameter with the classification. We
are interested in rules for networks classified as software-like and in the
accuracy of these rules, i.e., the proportion of networks that are correctly
classified. The rules found by 1R are shown in Table \ref{tab:rules}.

\begin{table}
\caption{Rules for predicting the classification of a synthetic network. S
stands for software-like and N stands for non software like.}
\centering
\begin{tabular}{|l|l|l|}
\hline
Model & Rule & Accuracy \\
\hline 
\hline
\multirow{2}{*}{BCR+}
     & $p_1 \ge 0.7 \Rightarrow S$ & \multirow{2}{*}{82.4\%}  \\ 
     & $p_1 < 0.7 \Rightarrow N$ & \\ 
\hline
\multirow{2}{*}{CGW}
     & $p_1 \ge 0.5 \Rightarrow S$ & \multirow{2}{*}{82.3\%} \\  
     & $p_1 < 0.5 \Rightarrow N$ & \\  
\hline
\multirow{2}{*}{LF}   
     & $\gamma \ge 2.44 \Rightarrow S$ & \multirow{2}{*}{78.9\%} \\ 
     & $\gamma < 2.44 \Rightarrow N$ & \\ 
\hline
\end{tabular}
\label{tab:rules}
\end{table}

The rules are very simple and, thus, easy to follow. Despite their simplicity,
they have high accuracy, approximately 80\% in all models.  
%These rules allow one to know if a particular combination of parameter values
%is likely to lead to software-like networks.

>>>>

%   ok: abstract
%   ok: say that we use DIRECTED GRAPHS
% TODO: review

% DEPENDENCY ANALYSIS
%
% component dependency isn't a good description...
% reference?
% usage?

% Martin Fowler
%
%   Coupling also occurs when code in one
%module uses code from another, perhaps by
%calling a function or accessing some data. 
%   When I use the term dependency, I use it as
%defined in the Unified Modeling Language
%(UML). So, the UI module depends on the do-
%main module if any code in the UI module ref-
%erences any code in the domain model - by
%calling a function, using some data, 
%or using types defined in the domain module
%
% A dependency is unidirectional

%A dependency is a semantic relationship where a change to the influent or
%independent modeling element may affect the semantics of the dependent
%modeling element. http://en.wikipedia.org/wiki/Dependency_(UML)

% Neeraj Sangal
% Dependencies are extracted from the code by a conventional static analysis
%
% Ao falar de dependências, cita
%   Designing Software for Ease of Extension and Contraction
%
%The LDM tool uses a standard notion of dependency, in which a
%module A depends on a module B if there are explicit references
%in A to syntactic elements of B. Currently, a module is a Java
%class, but a more fine-grained analysis is possible. This simple but
%effective notion of dependency works well for understanding
%design dependencies, in which modifications to one module might
%affect another. It is less well suited to determining runtime
%properties (such as how failures can propagate between modules),
%which require a deeper static analysis.
%
%---------
%
% Parnas
%
% A uses B if correct execution of B may be necessary for A to complete the
% task described in its specification.
%
% A uses B if there exist situations in which the correct functioning of A
% depends upon the availability of a correct implementation of B.
%
% "Uses" can also be formulated as "requires the presence of a correct version
% of".
%
% Uses vs. invokes

@@

% TODO: this paragraph can mostly disappear.
%
The time spent on the maintenance of a software system can be affected by the
assignment of developers to implementation components of the system.
If two development teams are responsible for maintaining components that are
highly interdependent, they need to coordinate their work often in order avoid
code duplication and breaking each other's component funcionality.

%One of the most obvious ways to reduce the time needed to develop or maintain a
%software system is to assign more developers to it. When combined with poor
%management, though, adding developers may in fact raise the development time,
%due to the increasing communication costs between developers \cite{Brooks1995}.
%
%The communication costs are affected by the assignment of developers to
%implementation components of a system.
%%The assignment of developers to implementation components of a system affects
%%the amount of communication needed between developers. 
%If two development teams are responsible for maintaining components that are
%highly interdependent, they need to coordinate their work often in order avoid
%code duplication and breaking each other's component funcionality.

Therefore, to optimize the development of a system, a manager can decompose the
system into modules, or groups of components, so that there are as few
inter-module dependencies as possible. % (i.e., the modules are weakly couple.).
%Each module is, then, threated as an unit of work assignment \cite{Parnas1972}:
Then, each module is assigned to one team and, because of the little inter-module
dependency, the communication between teams is reduced, therefore speeding up
the development.

To decompose a system into modules, a manager needs to have global knowledge of
the dependencies between its components. Although dependency analysis tools can
extract such information from the source code of the system, the task of finding
a good decomposition can be overwhelming because of the large number of
components and dependencies that need to be considered.

Software clustering algorithms, also known as architecture recovery algorithms
\cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005},
%\cite{Pollet2007}, 
automatically decompose a software system by analyzing the network of
dependencies between its components. %Although many algorithms have been
%proposed, few comparative studies exist.  empirical studies regarding the
%quality of the decompositions they find.  EVALUATION CRITERION
One important test for a software clustering algorithm is to apply it to a
collection of software systems with known reference decompositions made by
experienced developers \cite{Anquetil1999}. It is expected that a good algorithm
finds decompositions that are similar to reference decompositions, which can be
measured by a similarity metric such as MoJo \cite{Tzerpos1999} or EdgeSim
\cite{Mitchell2001}.

Unfortunately, there are few publicly available reference decompositions
\cite{Koschke2000} and, as a result, there are few studies that compare software
clustering algorithms. Also, because it is costly to obtain reference
decompositions for large systems, most studies test the algorithms with a couple
of small and medium systems \cite{Anquetil1999,Maqbool2007,Bittencourt2009}.

In our research we propose to evaluate software clustering algorithms by
applying them to synthetic, i.e., computer generated, component dependency
networks with built-in reference decompositions. This approach enables one to
test algorithms with arbitrarily large samples in a controlled manner.

A desirable property of synthetic networks is that they resemble their natural
counterparts. In this paper, we present three network models and show
empirically that all three models are capable of synthesizing networks that
resemble software networks. We consider object-oriented systems with classes as
components.

The paper is organized as follows: Section \ref{sec:models} presents three
network models.
%
Section \ref{sec:characterization} demonstrates a method to tell if a network
resembles a software network.
%
Section \ref{sec:evaluation} shows empirically that the three models synthesize
networks that resemble software networks.
%
Section \ref{sec:conclusion} discusses the impact of this research and future
work.

@@

@@

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, remarkable similarities have been
found in networks arising from domains as diverse as sociology, biology,
technology, and linguistics. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to class dependency networks, represented as
directed unweighted graphs, and it was shown that they are also scale-free
networks \cite{Myers2003,Valverde2002}. Thus, the study of the structure of
object-oriented software systems can benefit from the vast literature available
about scale-free networks.

In particular, many scale-free network models were proposed. These models are
algorithms that were proven, either formally or empirically, to synthesize
networks that are scale-free. In this section we present three models that
synthesize scale-free networks with built-in modular decompositions: BCR+, CGW,
and LF. The networks are represented as directed unweighted graphs with modules,
which makes them suitable for testing software clustering algorithms.

% directed unweighted graphs

%\subsection{ER (Erdos-Renyi)}
%\subsection{Configuration model} %XXX consider using with the most typical sw

@@

The algorithm 1R \cite{OneR} from machine learning was used to help discover
such relationship. It analyzes the parameters and the classification of each
network and finds a rule that relates the value of one single parameter with the
classification. Such rules can be evaluated according to their accuracy, i.e.,
the proportion of networks that are correctly classified. The rules found by 1R
are shown in Table \ref{tab:rules}.

\begin{table}
\caption{Rules for predicting the classification of a synthetic network. S
stands for software-like and N stands for non software like.}
\centering
\begin{tabular}{|l|l|l|}
\hline
Model & Rule & Accuracy \\
\hline 
\hline
\multirow{2}{*}{BCR+}
     & $p_1 \ge 0.7 \Rightarrow S$ & \multirow{2}{*}{82.4\%}  \\ 
     & $p_1 < 0.7 \Rightarrow N$ & \\ 
\hline
\multirow{2}{*}{CGW}
     & $p_1 \ge 0.5 \Rightarrow S$ & \multirow{2}{*}{82.3\%} \\  
     & $p_1 < 0.5 \Rightarrow N$ & \\  
\hline
\multirow{2}{*}{LF}   
     & $\gamma < 2.44 \Rightarrow S$ & \multirow{2}{*}{78.9\%} \\ 
     & $\gamma \ge 2.44 \Rightarrow N$ & \\ 
\hline
\end{tabular}
\label{tab:rules}
\end{table}

The rules are very simple and, thus, easy to follow. Despite their simplicity,
they have high accuracy, approximately 80\% for all models.  
%These rules allow one to know if a particular combination of parameter values
%is likely to lead to software-like networks.
